---
title: "Constructing Phylogenies with DECIPHER"
author: Aidan Lakshman^[ahl27@pitt.edu]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ConstructingPhylogenies}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coming Soon!
This tutorial will be coming soon! It will begin with the data from the previous step and cover the following topics:

* Phylogenetic reconstruction with `TreeLine()`
* Different input methods for `TreeLine()` (ex. `DistanceMatrix`)
* Different parameters for `TreeLine()` (ex. ML vs. MP, models, information criteria)

At the conclusion of this section, we will have constructed gene trees for each of the COGs identified in the previous section.

```{r eval=FALSE}
library(DECIPHER)

dna
ali <- AlignSeqs(dna)
DistanceMatrix(ali)

tree <- TreeLine(ali, maxTime=0)
Cophenetic(tree)

plot(tree)
```

```{css, echo=FALSE}
.pagination {
    display: flex;
    display: -webkit-flex;
    padding-left: 0;
    list-style: none;
    justify-content: center
}
```

<div class="center">
  <ul class="pagination pagination-lg">
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">&laquo;</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/Setup.html">1</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">2</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/GeneCallingAnnotation.html">3</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">4</a>
    </li>
    <li class="page-item active">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/ConstructingPhylogenies.html">5</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/CoevolutionNetworks.html">6</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/CoevolutionNetworks.html">&raquo;</a>
    </li>
  </ul>
</div>