---
title: "Gene Calling and Annotation with DECIPHER"
author: Aidan Lakshman^[ahl27@pitt.edu]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GeneCallingAnnotation}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Coming Soon!
This tutorial will be coming soon! It will begin with data from the previous step, and will cover the following topics:

* Finding protein coding regions in a genome with `FindGenes()`
* Finding non-coding RNAs in a genome with `FindNonCoding()`
* Classifying sequences using `IdTaxa()`

At the conclusion of this section, we will have identifying coding and non-coding regions in our input genomes along with predicted annotations of coding regions.

```{r eval=FALSE}
library(DECIPHER)
data("NonCodingRNA_Bacteria")
data("TrainingSet_16S")
dna
ncRNA <- NonCodingRNA_Bacteria

FindGenes(dna)
FindNonCoding(ncRNA, dna)

nogap <- RemoveGaps(dna)
ids <- IdTaxa(nogap, TrainingSet_16S)
```

```{css, echo=FALSE}
.pagination {
    display: flex;
    display: -webkit-flex;
    padding-left: 0;
    list-style: none;
    justify-content: center
}
```

<div class="center">
  <ul class="pagination pagination-lg">
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">&laquo;</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/Setup.html">1</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">2</a>
    </li>
    <li class="page-item active">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/GeneCallingAnnotation.html">3</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">4</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/ConstructingPhylogenies.html">5</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/CoevolutionNetworks.html">6</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">&raquo;</a>
    </li>
  </ul>
</div>