---
title: "Setup"
author: Aidan Lakshman^[ahl27@pitt.edu]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setup}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(downloadthis)
```

## Setting up your environment
This tutorial depends on a few key packages to work correctly. This page will outline
how to install all the necessary components for this workshop. If you're attending
this tutorial at Bioconductor 2022, [you can skip directly to the next page](https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html).

### Prerequisites
All content will be shown using the [R programming language](https://www.r-project.org/). 
For existing users, note that your version of `R` must be at least version `4.x.x` 
for compatibility with the packages we'll be using.

I also recommend using [RStudio](https://www.rstudio.com/products/rstudio/) as a 
development environment, since it's an excellent IDE with tons of features. 
This step is totally optional; feel free to use whatever development environment 
you prefer.

### Installing packages
This workshop depends on two main packages: `DECIPHER` and `SynExtend`. 
These will be installed via [Bioconductor](http://bioconductor.org/), a package 
manager for open source bioinformatics projects in R.

```{r eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install('DECIPHER')
BiocManager::install('SynExtend')
```

### Downloading Data
Click the below button to download all datafiles used in this workshop.

```{css echo=FALSE}
.hvr-grow-shadow {
  display: inline-block;
  vertical-align: middle;
  -webkit-transform: perspective(1px) translateZ(0);
  transform: perspective(1px) translateZ(0);
  box-shadow: 0 0 1px rgba(0, 0, 0, 0);
  -webkit-transition-duration: 0.3s;
  transition-duration: 0.3s;
  -webkit-transition-property: box-shadow, transform;
  transition-property: box-shadow, transform;
}
.hvr-grow-shadow:hover, .hvr-grow-shadow:focus, .hvr-grow-shadow:active {
  box-shadow: 0 10px 10px -10px rgba(0, 0, 0, 0.5);
  -webkit-transform: scale(1.1);
  transform: scale(1.1);
}
```
```{r echo=FALSE}
dir <- list.files(path=system.file('extdata', package='CompGenomicsBioc2022'),
                  full.names=TRUE)
downloadthis::download_file(
  path = dir,
  output_name = "exampleDataCGBioc2022",
  button_label = "Download example datafiles",
  button_type = "success",
  has_icon = TRUE,
  icon = "fa fa-download",
  self_contained = FALSE,
  class = "hvr-grow-shadow"
)
```

&nbsp;

&nbsp;

&nbsp;

&nbsp;


```{css, echo=FALSE}
.pagination {
    display: flex;
    display: -webkit-flex;
    padding-left: 0;
    list-style: none;
    justify-content: center
}
```

<div class="center">
  <ul class="pagination pagination-lg">
    <li class="page-item disabled">
      <a class="page-link" href="#">&laquo;</a>
    </li>
    <li class="page-item active">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/Setup.html">1</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">2</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/GeneCallingAnnotation.html">3</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">4</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/ConstructingPhylogenies.html">5</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/CoevolutionNetworks.html">6</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">&raquo;</a>
    </li>
  </ul>
</div>