---
title: "Setup"
author: Aidan Lakshman^[ahl27@pitt.edu]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setup}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up your environment
This tutorial depends on a few key packages to work correctly. This page will outline
how to install all the necessary components for this workshop. If you're attending
this tutorial at Bioconductor 2022, [you can skip directly to the next page](https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html).

### Prerequisites
All content will be shown using the [R programming language](https://www.r-project.org/). 
For existing users, note that your version of `R` must be at least version `4.x.x` 
for compatibility with the packages we'll be using.

I also recommend using [RStudio](https://www.rstudio.com/products/rstudio/) as a 
development environment, since it's an excellent IDE with tons of features. 
This step is totally optional; feel free to use whatever development environment 
you prefer.

### Installing packages
This workshop depends on two main packages: `DECIPHER` and `SynExtend`. 
These will be installed via [Bioconductor](http://bioconductor.org/), a package 
manager for open source bioinformatics projects in R.

```{r eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install('DECIPHER')
BiocManager::install('SynExtend')
```

### Downloading Data

Several toy datasets for the examples will be available in the associated GitHub 
repository for this workshop. 

```{css, echo=FALSE}
.pagination {
    display: flex;
    display: -webkit-flex;
    padding-left: 0;
    list-style: none;
    justify-content: center
}
```

<div class="center">
  <ul class="pagination pagination-lg">
    <li class="page-item disabled">
      <a class="page-link" href="#">&laquo;</a>
    </li>
    <li class="page-item active">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/Setup.html">1</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">2</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/GeneCallingAnnotation.html">3</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">4</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/ConstructingPhylogenies.html">5</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/CoevolutionNetworks.html">6</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">&raquo;</a>
    </li>
  </ul>
</div>