---
title: "Setup"
author: Aidan Lakshman^[ahl27@pitt.edu]
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setup}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up your environment
This tutorial depends on a few key packages to work correctly. This page will outline
how to install all the necessary components for this workshop. If you're attending
this tutorial at BioC2022, you can skip directly to the next page.

### Prerequisites
All content will be shown using the `R` programming language. You can install
`R` at [this website](https://www.r-project.org/) for your operating system. For 
existing users, note that your version of `R` must be at least version `4.x.x` 
for compatibility with the packages we'll be using.

I also recommend using RStudio as a development environment, since it's an 
excellent IDE with tons of features. This can be downloaded at [this link](https://www.rstudio.com/products/rstudio/). This step is totally optional;
feel free to use whatever development environment you prefer.

### Installing packages
This workshop depends on two main packages: `DECIPHER` and `SynExtend`. 
These will be installed via [Bioconductor](http://bioconductor.org/), a package 
manager for open source bioinformatics projects in R.

```{r eval=FALSE}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install('DECIPHER')
BiocManager::install('SynExtend')
```

### Downloading Data

Several toy datasets for the examples will be available in the associated GitHub 
repository for this workshop. 


<div class="center">
  <ul class="pagination pagination-lg">
    <li class="page-item disabled">
      <a class="page-link" href="#">&laquo;</a>
    </li>
    <li class="page-item active">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/Setup.html">1</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">2</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/GeneCallingAnnotation.html">3</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/FindingCOGs.html">4</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/ConstructingPhylogenies.html">5</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/CoevolutionNetworks.html">6</a>
    </li>
    <li class="page-item">
      <a class="page-link" href="https://www.ahl27.com/CompGenomicsBioc2022/articles/LoadingGenomeData.html">&raquo;</a>
    </li>
  </ul>
</div>